---
title: "Regression Models - Course Project"
author: "Lorena Zuniga"
output: pdf_document
fontsize: 9pt
---
#### Executive Summary
For this project, Motor Trend (a magazine about the automobile industry) is interested in exploring the relationship between a set of variables and miles per gallon (MPG). They need to answer the following questions:
1- Is an automatic or manual transmission better for MPG ?
2- Quantify the MPG difference between automatic and manual transmissions
This project use the mtcars dataset

#### Exploratory Analysis
Load the data:
```{r Echo =FALSE, results='hide'}
data("mtcars")
str(mtcars)
```
In the am column: 0 is automatic transmission, 1 = manual transmission.
We can observe that the MPG mean between each type of transmission is different:
```{r}
 autom<- mtcars[mtcars$am==0,]
 manual <-mtcars[mtcars$am==1,]
 c(mean(autom$mpg),mean(manual$mpg))
```
#### Inference
It is necessary to determine if the MPG mean difference is significant, then we can perform a t-test. Alpha value =0.05
```{r  results='hide'}
 t.test(autom$mpg,manual$mpg)
```
 The p-value is 0.001374, so in this case the difference between the MPG mean of the cars with manual transmission and the automatic cars is significant

#### Regression Analysis
At first we cant test a simple linear regression model using am as predictor and mpg as outcome

```{r  Echo =FALSE, results='hide'}
  summary (lm(mpg ~ am, data=mtcars))
```
The coefficient and intercept shows that manual transmission cars have 7.24 MPG and automatic transmission cars have 17.24 MPG.
The R-squared = 0.338 , that shows the model only explains 33.8% of the variance.

####Multivariable linear regression
In order to know which variables could be better to include in the multivariable linear regression model, we can get a correlation matrix, as follows:
```{r  Echo =FALSE, results='hide'}
  cor(mtcars)[1,]
```
From the resultant matrix we can say that wt and hp could be tested as predictors
```{r  Echo =FALSE, results='hide'}
  simpleModel <- lm(mpg ~ am,data=mtcars)
  multiModel <- lm(mpg ~ am + hp + wt, data=mtcars)
```
Comparing both models with ANOVA:
```{r Echo =FALSE, results='hide'}
  anova(simpleModel, multiModel)
```
The p-value is 3.7e-09 and therefore the difference between both models is significant

```{r Echo =FALSE}
 summary(multiModel)  
```
Observing the coefficients, it is possible to say that, in average, cars with manual transmission have 2.083MPG more than cars with automatic transmission.
For the multivariable model the R-squared = 0.84, it is possible to say that the multivariable model explain the  84% of the variance.

#### Residuals
 It is verified that the mean of the residuals is near 0:
````{r}
  mean(multiModel$residuals)
````
 Also it is possible verify that each one of the MPG predictors (hp, wt and am) are uncorrelated with the residuals
The residuals chart can be found in the Apendix
```{r}
  cov(multiModel$residuals,mtcars$hp)
  cov(multiModel$residuals,mtcars$wt)
  cov(multiModel$residuals,mtcars$am)
````
#### Apendix
```{r}
   par(mfrow=c(2,2))
   par(mar=rep(2,4))
   plot(multiModel)
```
The complete code for this project can be found here: (https://github.com/lorenazs/regressionmodels/blob/master/Project/project.Rmd) 